---
title: "p8105_hw1_ryd2106"
author: "Risa Dickerman"
date: "2026-09-25"
output: github_document
---


```{r set up, include = FALSE}
#Library and data set load in
library(tidyverse)

#Hide annoying error messages
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Problem 1
```{r dataframe}
#Load in data frame
data("penguins", package = "palmerpenguins")

#Calculate the mean flipper length per species
flip_len_spec = aggregate(penguins$flipper_length_mm,
                    list(penguins$species), mean, na.rm=TRUE)

```
The penguins data set from the palmerpenguins package described the morphological features of `r nrow(penguins)` different birds. There were `r ncol(penguins)` different variables observed. These variables included the bird's `r colnames(penguins[3:7])` in addition to the penguin's `r colnames(penguins[1])`, the `r colnames(penguins[2])` they lived on, and the `r colnames(penguins[8])` the samples were collected. The species, island, and sex variables were all factorial (category) values denoted by fct. The bill_length_mm and bill_depth_mm variables were numeric (dbl) while the year, flipper_length_mm, and body_mass_g were all integers (int).

This study found several interesting statistics of the penguins on the following islands: `r unique(penguins$island)`. For example, it was found that the mean flipper length of all these birds lie at `r signif(mean(penguins$flipper_length_mm, na.rm=TRUE), digits=3)`mm. The mean flipper lengths of the `r flip_len_spec[1,1]`, `r flip_len_spec[2,1]`, and `r flip_len_spec[3,1]` penguins were `r signif(flip_len_spec[1,2], digits=3)`mm, `r signif(flip_len_spec[2,2], digits = 3)`mm, and `r signif(flip_len_spec[3,2], digits=3)`mm respectively.

Visualizations of certain features are presented below:

A scatter plot of the flipper length vs the bill length of these penguins is shown below:
```{r scatter_plot}
#Create the scatterplot and save
ggplot(data = penguins, mapping = aes(x = flipper_length_mm, y = bill_length_mm)) + geom_point() + geom_point(aes(color=species))
ggsave("hw_scatterplot.pdf")
```

#Problem 2

```{r data_frame}
#Setting the seed
set.seed(21061)

#Producing a data frame
prob2_df = tibble(
  vec_random = rnorm(10, mean = 0, sd= 1 ),
  vec_logical = (vec_random > 0),
  vec_character = c("It", "was", "the", "best", "of", "times,", "it", "was", "the", "worst.."),
  vec_factorial = factor(c("mild", "medium", "medium", "hot", "mild", "hot", "medium", "hot", "medium", "mild"))
)

#Take the mean of the variables
mean_random = mean(
  pull(prob2_df,1))
mean_logical = mean(
  pull(prob2_df,2))
mean_character = mean(
  pull(prob2_df,3))
mean_factorial =mean(
  pull(prob2_df,4))
prob2_df
```

When taking the mean of each variable in the prob2_df, some make sense while others do not. The most logical result is the random vector's (vec_randomn) mean of `r mean_random`. The mean of the logical vector (vec_logical), however, returns the numerical value `r mean_logical` even though it is taken from the logical values TRUE and FALSE. Finally, taking the mean of both the character (vec_character) and factorial (vec_factorial) vectors return a `r mean_factorial` value.
