---
title: "p8105_hw1_ryd2106"
output: html_document
date: "2026-09-23"
---

```{r setup, include=FALSE}
library(tidyverse)
```

#Problem 1

```{r load in penguins}
data("penguins", package = "palmerpenguins")
```

This data set describes morphological features of `r nrow(penguins)` different penguins.
`r ncol(penguins)` different variables were observed. These included the bird's `r colnames(penguins[3:7])` in addition to the `r colnames(penguins[1])` of the bird, the `r colnames(penguins[2])` they lived on, and the `r colnames(penguins[8])` the samples were collected.

This study found several interesting statistics of the penguins on the following islands: `r unique(penguins$island)`. For example it was found that the mean flipper length of these birds lie at `r signif(mean(penguins$flipper_length_mm, na.rm=TRUE), digits=3)` milimeters. The largest bird observed weighed in at `r max(penguins[6], na.rm=TRUE)` grams while the smallest bird weighed `r min(penguins[6], na.rm=TRUE)` grams.


the data in this dataset, including names / values of important variables
the size of the dataset (using nrow and ncol)
the mean flipper length


```{r }

```



```{r }

```

