---
title: "p8105_hw1_ryd2106"
output: html_document
date: "2026-09-23"
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Problem 1

```{r load in penguins, include=FALSE}
data("penguins", package = "palmerpenguins")
flip_len_spec = aggregate(penguins$flipper_length_mm,list(penguins$species), mean, na.rm=TRUE)
```

Th penguins data set from the palmerpenguins package describes morphological features of `r nrow(penguins)` different penguins.
`r ncol(penguins)` different variables were observed. These included the bird's `r colnames(penguins[3:7])` in addition to the `r colnames(penguins[1])` of the bird, the `r colnames(penguins[2])` they lived on, and the `r colnames(penguins[8])` the samples were collected.

This study found several interesting statistics of the penguins on the following islands: `r unique(penguins$island)`. For example it was found that the mean flipper length of these birds lie at `r signif(mean(penguins$flipper_length_mm, na.rm=TRUE), digits=3)`mm. The flipper length of the `r flip_len_spec[1,1]` penguin had a mean of `r signif(flip_len_spec[1,2], digits=3)`mm, the `r flip_len_spec[2,1]` species mean was `r signif(flip_len_spec[2,2], digits = 3)`mm, while the `r flip_len_spec[3,1]`'s mean was `r signif(flip_len_spec[3,2], digits=3)`mm.


Visualizations of certain features are presented below. 

Flipper length was compared to bill length for each species.
```{r }
flip_len_spec = aggregate(penguins$flipper_length_mm,list(penguins$species), mean, na.rm=TRUE)
flip_len_spec
flip_len_spec[3,3]
```



```{r }

```

